Towards Automated Merging of Code Clones in Object-Oriented Programming Languages\\- Work in Progress -

### 2019-05-06 sattose/paper.tex
Sander and Xander

\section{Redefining type 1, 2 and type 3 clones}\label{chap:clonetypes}
In this chapter, we will propose solutions for shortcomings in type 1, type 2 and type 3 clones to increase the chance that we can merge the clone while improving the design. Type-4 clones are, because of the serious challenges involved in their detection and refactoring, out of the scope for this study.

\subsection{Type 1 clones} \label{chap:type1clones}
Type 1 clones are identical clone fragments except for variations in whitespace and comments. However, when two clone fragments are textually identical, it does not yet indicate that they are actually identical. Even when textually equal, method calls can refer to different methods, type declarations can refer to different types and even variables can be of a different type. This could invalidate a refactoring opportunity for such a code fragment. Because of that, we recommend redefining type-1 clones for refactoring. We propose to:
\begin{itemize}
  \item \textbf{Compare the equality of the fully qualified method signature for method references.} If an identifier is fully qualified, it means we specify the full location of its declaration (e.g. \texttt{com.simonbaars.fruitgame.Apple} for an \texttt{Apple} object). This way we can validate whether two method references, like method calls, are actually equal. In the method signature, not only the fully qualified identifier of the method should be considered, but also the type of all its arguments. This way we can be sure that two potentially cloned method references do not point to overloaded variants (in a case that the data type of arguments is overloaded).
  \item \textbf{Compare the equality of the fully qualified identifier for type references.} This way we can be sure that two referenced types are actually equal, and that they are not just two types with the same name.
  \item \textbf{Compare the equality the fully qualified identifier for variable usages.} Two cloned lines might use a variable with the same name, but different types. This might pose serious challenges on refactoring, as the variables might not concern the same object or primitive. To check this, we need to track the declaration of variables and from this infer the fully qualified identifier of its type.
\end{itemize}

@@

Because of that, for this research, we have looked into a redefinition of type 2 clones to be able to detect such clones that can and should be merged. Our definition ensures functional similarity by applying the following changes to type 2 clones:

@@

However, not always can a clone easily be merged. Sometimes a clone spans a statement partially (like a for-loop of which only it's declaration and a part of the body is cloned). Merging the clones can be harder in such instances. Also, the cloned code can contain a complex control structure, like labels, return, break, continue, etc. In such instances, more conditions apply to be able to conduct a refactoring, if advisable at all.

@@

\subsection{Next steps}
The next step will be to integrate our redefinitions of type 1, 2 and 3 clones into CloneRefactor. This way, we can run all scripts that we used for our measurements for these novel clone definitions. We are curious to see how it compares with the results we had so far.

As we want to refactor duplicates fully automated, it is very important that both code fragments have exactly the same functionality. Because of this, we want to look into building a SymbolTable of projects we scan and their dependencies. This allows us to further assess the equality of code fragments. Currently, we compare fragments, as most clone detection tools do, on basis of static analysis of only the target project. However, because of this, we omit valuable information like fully qualified identifiers for types. By parsing not only a project, but also its dependencies, we can better assess whether two code fragments are actually equal, or whether they are calling different methods but with the same type signature (we are currently limited in distinguishing these).

>>>>

\section{Addressing and solving problems with clone type definitions}\label{chap:clonetypes}
In this chapter, we will propose solutions for shortcomings in type 1, type 2 and type 3 clones \cite{roy2007survey} to increase the chance that we can merge the clone while improving the design. Type-4 clones are, because of the serious challenges involved in their detection and refactoring, out of the scope for this study.

\subsection{Type 1 clones} \label{chap:type1clones}
Type 1 clones are identical clone fragments except for variations in whitespace and comments. However, when two clone fragments are textually identical, it does not yet indicate that they are actually identical. Even when textually equal, method calls can refer to different methods, type declarations can refer to different types and even variables can be of a different type. This could invalidate a refactoring opportunity for such a code fragment. Because of that, we propose to:
\begin{itemize}
  \item \textbf{Compare the equality of the fully qualified method signature for method references.} If an identifier is fully qualified, it means we specify the full location of its declaration (e.g. \texttt{com.simonbaars.fruitgame.Apple} for an \texttt{Apple} object). This way we can validate whether two method references, like method calls, are actually equal. In the method signature, not only the fully qualified identifier of the method should be considered, but also the type of all its arguments. This way we can be sure that two potentially cloned method references do not point to overloaded variants (in a case that the data type of arguments is overloaded).
  \item \textbf{Compare the equality of the fully qualified identifier for type references.} This way we can be sure that two referenced types are actually equal, and that they are not just two types with the same name.
  \item \textbf{Compare the equality the fully qualified identifier for variable usages.} Two cloned lines might use a variable with the same name, but different types. This might pose serious challenges on refactoring, as the variables might not concern the same object or primitive. To check this, we need to track the declaration of variables and from this infer the fully qualified identifier of its type.
\end{itemize}

@@

Because of that, for this research, we have looked solving these problems with type 2 clones to be able to detect such clones that can and should be merged. Our definition ensures functional similarity by applying the following changes to type 2 clones:

@@

However, not always can a clone easily be merged. Sometimes a clone spans a statement partially (like a for-loop of which only it's declaration and a part of the body is cloned). Merging the clones can be harder in such instances. Also, the cloned code can contain statements like labels, return, break, continue, etc. In such instances, more conditions may apply to be able to conduct a refactoring, if advisable at all.

@@

\subsection{Next steps}
The next step will be to integrate our solutions for the problems with the clone types, as explained in section \ref{chap:clonetypes}, into our CloneRefactor tool. This way, we can run all scripts that we used for our measurements for these novel clone definitions. We are curious to see how it compares with the results we had so far.
